---
title: "Exploration One Acre Fund Plot level data for maize yield prediction"
author: "Benson Kenduiywo"
date: "01/12/2020"
output: html_document
---

```{r, global_options, tidy = TRUE,echo=TRUE}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60))

```

# Introduction

load the acre fund data plot level data and MODIS data

```{r d1}
rm(list=ls(all=TRUE))
library(raster)
library(data.table)
f <- "D:/UCDavis/QUIIC/github/oaf_yield/oaf/KEN_oaf_8day_MODIS_vi_006_250m.csv"
d <- fread(f)
tail(d)
y <- substring(d$dates, 1,4)
table(y)

shp <- shapefile("../oaf/kenya_crop_yield_oaf_2015_18.shp")
```

Merge remote sensing information to corresponding plot information i.e.:

|Column|Description|
|---|---|
|uid|unique plot/field/farmer identifier to merge with GEE export data| 
|Year|year of data collection|
|Season|season of data collection|
|Country|country of data collection|
|yld_kg_|shelled maize grain yield in kg/ha|
|PltSz_h|plot size in hectares|
|CAN_kg_|CAN (topdress fertilizer) in kg/ha|
|DAP_kg_|DAP (basal fertilizer) applied at planting) in kg/ha|
|NPK_kg_|NPK in kg/ha|
|Ur_kg_h|urea (topdress fertilizer) in kg/ha|
|Variety|variety of maize planted, if known|
|PlntngD|date of planting|

Extract MODIS data corresponding to plot level years(2015--2017)

```{r d2}
i <- unique(shp$Year)
df1 <- d[y %in% i,]
```

Aggregate plot level satellite signal per year per season (March -- September, in this case only Long Rain (LR) season only). To compute daily average signal per season in each year, we first Create a function to compute means for LR season.

```{r fxn1}
seasonMean <- function(year, df, seasons=1:2) {
  res <- list()
  for (i in seasons) {
    season <- ifelse(i==1, "long", "short")
    if (season =="long") {
      sdate <- paste0(year, "-03-01")
      edate <- paste0(year, "-09-30")
      season <- "LR"
    } else if (season =="short") {
      sdate <- paste0(year-1, "-10-01")
      edate <- paste0(year, "-02-28")
      season <- "SR"
    } else {
      stop("Define season")
    }
    ydf <- df[df$dates >= sdate & df$dates <= edate, ]
    ym <- aggregate(ydf[,1], ydf[,2, drop=FALSE], mean, na.rm=T)
    ym$Year <- year
    ym$season <- season
    res[[i]] <- ym    
  }  
  do.call(rbind, res)  
}

```


Compute season mean satellite signal.

```{r d3}
df1$dates <- as.Date(df1$dates, format = "X%Y%m%d")
years <- unique(format(df1$dates, format="%Y") )
#EVI
e <- df1[,c('EVI','uid',"dates")]
temp <- lapply(years, seasonMean, e, seasons=1)
e <- do.call(rbind, temp)
#NDVI
n <- df1[,c('NDVI','uid',"dates")]
temp <- lapply(years, seasonMean, n, seasons=1)
n <- do.call(rbind, temp)
df2 <- cbind(e[,c('uid','EVI')],n[,c('NDVI','Year','season')])
```

Merge plot level yield data with MODIS data. 

```{r d4}
names(shp)
names(df2)
df3 <- merge(df2, shp, by=c("uid","uid", "Year"))

```

Explore the relationship between yield and satellite signal.

```{r d5}
x11()
plot(df3$EVI, df3$yld_kg_, ylab="Maize yield (kg/ha)", xlab='EVI')
x11()
plot(df3$NDVI, df3$yld_kg_, ylab="Maize yield (kg/ha)", xlab='NDVI')

```
 
NOTE: Some of the yield per ha are higher than normal....On average in Kenya yields per ha range between 4--8 tonnes per hectare. 
 

Let's fit an ordinary regression model.

```{r lm1}
#EVI
lm1 <- lm(df3$yld_kg_~df3$EVI)
summary(lm1)
#NDVI
lm2 <- lm(df3$yld_kg_~df3$NDVI)
summary(lm2)

```

Model fitting is not good.

 


